---
title: "survey analysis plots"
output: html_document
date: "2025-05-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Load packages for data handling, visualization, and formatting
library(tidyverse)
library(here)
library(readxl)
library(stringr)
library(gt)
library(openxlsx)
library(survival)
library(survminer)
library(psych)
library(viridis)
library(ggtext)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(tidytext)

# Source custom functions
source("functions.R")

# Set global options
options(stringsAsFactors = FALSE)
options(scipen = 100, digits = 4)


# Load reference data: country-to-region mapping
country_dict <- readxl::read_excel("data/countries_regions.xlsx")

file_path_plots <- "figures"

# Define path to directory containing raw survey CSVs
directory_path <- "data"

# Identify all CSV files in the directory
csv_files <- list.files(path = directory_path, pattern = "\\.csv$")

# Load each CSV and store in a list of data frames
data_list <- lapply(csv_files, function(file) {
  read_csv(file.path(directory_path, file))
})

# Combine all individual survey CSVs into one data frame
df <- bind_rows(data_list)

df[df == ""] <- NA
df[df == "#N/A"] <- NA
df[df == "Other. Please specify:"] <- "Other"
df[df == "None"] <- NA

# Remove metadata rows (first two) from the combined data frame
df <- df[-c(1, 2), ]

df <- merge(df, country_dict, by.x = "Q8", by.y = "name", all.x = TRUE)

responses_n <- nrow(df)

#if Q1 was not answered, we can assume the person exited the survey without answering questions
drop_outs2 <- df %>% filter(is.na(Q1))
drop_outs_sum <- nrow(drop_outs2)

df <- df %>% filter(!is.na(Q1))
df <- df %>%
  filter(!is.na(df$StartDate))

```

Calculate denominators by survey section 

```{r include=FALSE}
# Compute completion status for Metrics section (at least 2 responses)
sub_df_metrics <- df[,c(9,56:94)] 

filtered_df <- sub_df_metrics %>%
  mutate(metrics_section_answered = rowSums(!is.na(across(-ResponseId)))) %>%
  filter(metrics_section_answered >= 2) 

filtered_df$metrics_section_answered <- ifelse(filtered_df$metrics_section_answered > 0, 1, 0)

metrics_section <- filtered_df[,c("ResponseId", "metrics_section_answered")]

# Compute completion status for Questions section (at least 2 responses)
sub_df_questions <- df[,c(9,95:160)]

filtered_df <- sub_df_questions %>%
  mutate(questions_section_answered = rowSums(!is.na(across(-ResponseId)))) %>%
  filter(questions_section_answered >= 2) 

filtered_df$questions_section_answered <- ifelse(filtered_df$questions_section_answered > 0, 1, 0)

questions_section <- filtered_df[,c("ResponseId", 
                                          "questions_section_answered")]

# Compute completion status for Evaluation section (at least 2 responses)
sub_df_evaluation <- df[,c(9,161:164)]

filtered_df <- sub_df_evaluation %>%
  mutate(evaluation_section_answered = rowSums(!is.na(across(-ResponseId)))) %>%
  filter(evaluation_section_answered >= 2) 

filtered_df$evaluation_section_answered <- ifelse(filtered_df$evaluation_section_answered > 0, 1, 0)

evaluation_section <- filtered_df[,c("ResponseId", 
                                            "evaluation_section_answered")]

#Confidence section
sub_df_confidence <- df[,c(9,165:174)]

filtered_df <- sub_df_confidence %>%
  mutate(confidence_section_answered = rowSums(!is.na(across(-ResponseId)))) %>%
  filter(confidence_section_answered >= 1) 

filtered_df$confidence_section_answered <- ifelse(filtered_df$confidence_section_answered > 0, 1, 0)

confidence_section <- filtered_df[,c("ResponseId", 
                                            "confidence_section_answered")]

#Barriers section
sub_df_barriers <- df[,c(9,175:186)]

filtered_df <- sub_df_barriers %>%
  mutate(barriers_section_answered = rowSums(!is.na(across(-ResponseId)))) %>%
  filter(barriers_section_answered >= 2) 

filtered_df$barriers_section_answered <- ifelse(filtered_df$barriers_section_answered > 0, 1, 0)
barriers_section <- filtered_df[,c("ResponseId", 
                                        "barriers_section_answered")]

```

Remove empty survey responses & aggregate policy engagement

```{r include=FALSE}
df <- df %>%
  left_join(metrics_section, by = "ResponseId") %>%
  left_join(questions_section, by = "ResponseId") %>%
  left_join(evaluation_section, by = "ResponseId") %>%
  left_join(barriers_section, by = "ResponseId")%>%
  left_join(confidence_section, by = "ResponseId")

# anyone who has not contributed to any policy dialogues around COVID-19 are excluded here
dropout <- nrow(df[(df$Q2 == "No" & df$Q4 == "No"), ])
df <- df[!(df$Q2 == "No" & df$Q4 == "No"), ] 

# summarizing Q2 and Q4 (due to many levels in diagram) - either they engaged in policy dialogues themselves or their supervisor did (Q4)

df$Q3_national <- ifelse(!is.na(df$Q3_1) | 
                              !is.na(df$Q3_7) | 
                              !is.na(df$Q3_2) | 
                              !is.na(df$Q3_6) | 
                              !is.na(df$Q3_3) | 
                              !is.na(df$Q3_4) |
                              !is.na(df$Q5_1) |
                              !is.na(df$Q5_2) |
                              !is.na(df$Q5_7) |
                              !is.na(df$Q5_6) |
                              !is.na(df$Q5_4),1,NA )

df$Q3_sub_national <- ifelse(!is.na(df$Q3_5) | 
                              !is.na(df$Q3_8) | 
                              !is.na(df$Q3_10)|
                              !is.na(df$Q5_5) |
                              !is.na(df$Q5_8) |
                              !is.na(df$Q5_10),1,NA )

df$Q3_field <- ifelse(!is.na(df$Q3_11) | 
                              !is.na(df$Q3_12),1,NA )

df$Q3_other <- ifelse(!is.na(df$Q3_16) | 
                              !is.na(df$Q3_15),1,NA )

# remove columns after aggregation
columns_to_remove <- c("Q3_1", "Q3_7", "Q3_2", "Q3_6", "Q3_3", "Q3_4",
                       "Q5_1", "Q5_2", "Q5_7", "Q5_6", "Q5_4",
                       "Q3_5", "Q3_8", "Q3_10", "Q5_5", "Q5_8", "Q5_10",
                       "Q3_11", "Q3_12","UserLanguage", "DistributionChannel",
                       "Q0","Status", "EndDate")

df <- df[ , -which(names(df) %in% columns_to_remove)] ## works as expected

df <- reorder_factor_levels(df, "Q1", "Other") 

df$q1_group <- ifelse(df$Q1 == "University/Research institution", "University/Research institution", "Other")

df <- df %>%
  mutate(Q1 = str_wrap(Q1, width = 20))

#relocate columns 
df <- df %>% 
  relocate(Q10_4, .before = Q8) %>% 
  relocate(Q10_5, .after = Q10_4) %>% 
  relocate(Q10_6, .after = Q10_5) %>% 
  relocate(income_group, .after = Q8) %>%
  relocate(metrics_section_answered, .after = income_group)  %>%
  relocate(questions_section_answered, .after = metrics_section_answered)%>%
  relocate(evaluation_section_answered, .after = questions_section_answered)%>%
  relocate(confidence_section_answered, .after = evaluation_section_answered)%>%
  relocate(barriers_section_answered, .after = confidence_section_answered) %>%
  relocate(Q3_national, .after = barriers_section_answered)%>%
  relocate(Q3_sub_national, .after = Q3_national)%>%
  relocate(Q3_field, .after = Q3_sub_national)%>%
  relocate(Q3_other, .after = Q3_field)

df$Q14_4 <- ifelse(df$Q14_4 == "Single number / Point estimate", "Point estimates", df$Q14_4)

df <- df %>%
  filter(!is.na(df$ResponseId))

```

Supplement 3: Survival curve to show dropout

```{r echo=FALSE}

# Clean up income_group
df$income_group <- ifelse(is.na(df$income_group), 
                          "Non-country-specific", 
                          df$income_group)

# Create a trimmed dataset
df_survival <- df[, c("ResponseId",
                      "income_group",
                      "metrics_section_answered",
                      "barriers_section_answered",
                      "questions_section_answered",
                      "evaluation_section_answered",
                      "confidence_section_answered",
                      "q1_group")]

# Define prettier labels for the sections
label_map <- c(
  metrics_section_answered = "Metrics",
  barriers_section_answered = "Barriers",
  questions_section_answered = "Questions",
  evaluation_section_answered = "Evaluation",
  confidence_section_answered = "Confidence"
)

# Set ordered factor levels for income group
df_survival$income_group <- factor(df_survival$income_group,
                                   levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"))

# Get total number of 1s per section
section_counts <- df_survival %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "section", values_to = "count")

# Order sections by descending count
section_counts <- section_counts %>%
  arrange(desc(count))

ordered_sections <- section_counts$section

# Sum numeric columns by income group
grouped_sum <- df_survival %>%
  group_by(income_group) %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE))

# Convert to long format for plotting
grouped_sum_long <- grouped_sum %>%
  pivot_longer(-income_group, names_to = "section", values_to = "value")

# Apply descending section order to both datasets
grouped_sum_long$section <- factor(grouped_sum_long$section, levels = ordered_sections)
section_counts$section <- factor(section_counts$section, levels = ordered_sections)

income_group_colors <- c(
  "HIC" = "#66c2a5",
  "UMIC" = "#fc8d62",
  "LMIC" = "#8da0cb",
  "LIC" = "#e78ac3",
  "Non-country-specific" = "#a6d854"
)

# Plot
ggplot(grouped_sum_long, aes(x = section, y = value, color = income_group, group = income_group)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  geom_text(data = section_counts,
            aes(x = section, y = 46, label = paste0("n = ", count)),
            inherit.aes = FALSE,
            size = 4,
            vjust = -0.5) +
  scale_x_discrete(labels = label_map) +
  scale_color_manual(values = income_group_colors) +  # ✅ consistent colors
  theme_minimal() +
  labs(
    title = "",
    x = "Survey Section",
    y = "Number of Respondents",
    color = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 0),
    legend.position = "bottom"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

ggsave(file.path(file_path_plots, "survey_dropout.png"), height = 6, width = 7, units = "in")

```


Figure 2: Differences in which format evidence communicated and Received by income group

```{r echo=FALSE}

metrics_counts <- df %>%
  filter(metrics_section_answered > 0) %>%
  group_by(income_group)  %>%
  summarise(respondents = n())

Received_format_overall <- summarize_columns_count(
  df = df,
  group_col = "income_group", 
  summary_cols = c("Q13_1", "Q13_2", "Q13_6", 
                            "Q13_4", "Q13_3", "Q13_5", 
                            "Q13_7"), 
  section = "metrics_section_answered")

names(Received_format_overall) <- gsub("\\(.*?\\)|\\[.*?\\]", "", names(Received_format_overall))
names(Received_format_overall) <- gsub("\\s$", "", names(Received_format_overall)) 
names(Received_format_overall) <- gsub("\\)$", "", names(Received_format_overall))

# Reshape the data from wide to long format
df_long <- Received_format_overall %>% 
  pivot_longer(cols = c("Map":"Scientific research paper"), 
               names_to = "format", 
               values_to = "count")

df_long_Received <- df_long
df_long_Received$type <- "Received"

Preferred_format_overall <- summarize_columns_count(
  df = df,
  group_col = "income_group",
  summary_cols = c("Q14_1", "Q14_2", "Q14_6", 
                            "Q14_4", "Q14_3", "Q14_5", 
                            "Q14_7"), 
  section = "metrics_section_answered")

names(Preferred_format_overall) <- gsub("\\(.*?\\)|\\[.*?\\]", "", names(Preferred_format_overall))
names(Preferred_format_overall) <- gsub("\\s$", "", names(Preferred_format_overall)) 
names(Preferred_format_overall) <- gsub("\\)$", "", names(Preferred_format_overall))
names(Preferred_format_overall)[6] <- c("Point estimates")

# Reshape the data from wide to long format
df_long <- Preferred_format_overall %>% 
  pivot_longer(cols = c("Map":"Scientific research paper"), 
               names_to = "format", 
               values_to = "count")

df_long_Preferred <- df_long
df_long_Preferred$type <- "Preferred"

df_long <- rbind(df_long_Received, df_long_Preferred)

df_long$type <- factor(df_long$type, levels=c("Received","Preferred"))

df_long <- df_long %>%
  mutate(format = str_wrap(format, width = 20))

df_long <- df_long %>%
  left_join(metrics_counts, by = "income_group")

# Set income_group factor order
df_long$income_group <- factor(df_long$income_group, levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"))

# Create ordered income_group_label based on the ordered income_group
df_long <- df_long %>%
  mutate(income_group_label = paste0(as.character(income_group), " (n = ", respondents, ")"))

# Ensure consistent label order in facets
income_levels_ordered <- df_long %>%
  distinct(income_group, income_group_label) %>%
  arrange(factor(income_group, levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"))) %>%
  pull(income_group_label)

df_long$income_group_label <- factor(df_long$income_group_label, levels = income_levels_ordered)

# Prepare data
df_long <- df_long %>%
  mutate(
    proportion = count / respondents,
    format = fct_rev(factor(format)),  # so it appears top-down on y-axis
    type = factor(type, levels = c("Received", "Preferred")),  # order L to R
    income_group_label = factor(income_group_label, levels = income_levels_ordered)
  )

# Step 1: Convert format to numeric base position
df_long <- df_long %>%
  mutate(format = fct_rev(factor(format))) %>%  # make top-down order
  mutate(format_num = as.numeric(format))

# Step 2: Add small offset per income group
income_offsets <- tibble(
  income_group_label = unique(df_long$income_group_label),
  offset = seq(-0.3, 0.3, length.out = length(unique(df_long$income_group_label)))
)

df_long <- df_long %>%
  left_join(income_offsets, by = "income_group_label") %>%
  mutate(y = format_num + offset)

ordered_income_groups <- c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific")
group_sizes <- df_long %>%
  select(income_group, income_group_label) %>%
  distinct()

ordered_labels <- group_sizes$income_group_label[match(ordered_income_groups, group_sizes$income_group)]

fill_colors <- setNames(RColorBrewer::brewer.pal(n = length(ordered_labels), "Set2"), ordered_labels)

ggplot(df_long, aes(x = proportion, y = y, group = interaction(income_group_label, format), color = income_group_label)) +
  geom_line(size = 1) +
  geom_point(aes(shape = type), size = 3) +
  scale_y_continuous(
    breaks = 1:length(levels(df_long$format)),
    labels = levels(df_long$format)
  ) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_color_manual(values = fill_colors) + 
  labs(x = "Proportion of Respondents", y = "Format Type", color = "Income Group", shape = "") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 14),        # Increased X title
    axis.title.y = element_text(size = 14),        # Increased Y title
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),         # Larger legend text
    legend.title = element_text(size = 13),        # Larger legend title
    panel.grid.major.y = element_line(color = "grey90")
  )

ggsave(file.path(file_path_plots, "Received_Preferred_by_income.png"), width = 14, height = 10)

```

```{r}

metrics_counts <- df %>%
  filter(metrics_section_answered > 0) %>%
  group_by(income_group)  %>%
  summarise(respondents = n())

Received_format_overall <- summarize_columns_count(
  df = df,
  group_col = "income_group", 
  summary_cols = c("Q13_1", "Q13_2", "Q13_6", 
                            "Q13_4", "Q13_3", "Q13_5", 
                            "Q13_7"), 
  section = "metrics_section_answered")

names(Received_format_overall) <- gsub("\\(.*?\\)|\\[.*?\\]", "", names(Received_format_overall))
names(Received_format_overall) <- gsub("\\s$", "", names(Received_format_overall)) 
names(Received_format_overall) <- gsub("\\)$", "", names(Received_format_overall))

# Reshape the data from wide to long format
df_long <- Received_format_overall %>% 
  pivot_longer(cols = c("Map":"Scientific research paper"), 
               names_to = "format", 
               values_to = "count")

df_long_Received <- df_long
df_long_Received$type <- "Received"

Preferred_format_overall <- summarize_columns_count(
  df = df,
  group_col = "income_group",
  summary_cols = c("Q14_1", "Q14_2", "Q14_6", 
                            "Q14_4", "Q14_3", "Q14_5", 
                            "Q14_7"), 
  section = "metrics_section_answered")

names(Preferred_format_overall) <- gsub("\\(.*?\\)|\\[.*?\\]", "", names(Preferred_format_overall))
names(Preferred_format_overall) <- gsub("\\s$", "", names(Preferred_format_overall)) 
names(Preferred_format_overall) <- gsub("\\)$", "", names(Preferred_format_overall))
names(Preferred_format_overall)[6] <- c("Point estimates")

# Reshape the data from wide to long format
df_long <- Preferred_format_overall %>% 
  pivot_longer(cols = c("Map":"Scientific research paper"), 
               names_to = "format", 
               values_to = "count")

df_long_Preferred <- df_long
df_long_Preferred$type <- "Preferred"

df_long <- rbind(df_long_Received, df_long_Preferred)

df_long$type <- factor(df_long$type, levels=c("Received","Preferred"))

df_long <- df_long %>%
  mutate(format = str_wrap(format, width = 20))

df_long <- df_long %>%
  left_join(metrics_counts, by = "income_group")

# Set income_group factor order
df_long$income_group <- factor(df_long$income_group, levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"))

# Create ordered income_group_label based on the ordered income_group
df_long <- df_long %>%
  mutate(income_group_label = paste0(as.character(income_group), " (n = ", respondents, ")"))

# Ensure consistent label order in facets
income_levels_ordered <- df_long %>%
  distinct(income_group, income_group_label) %>%
  arrange(factor(income_group, levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"))) %>%
  pull(income_group_label)

df_long$income_group_label <- factor(df_long$income_group_label, levels = income_levels_ordered)

# Prepare data
df_long <- df_long %>%
  mutate(
    proportion = count / respondents,
    format = fct_rev(factor(format)),  # so it appears top-down on y-axis
    type = factor(type, levels = c("Received", "Preferred")),  # order L to R
    income_group_label = factor(income_group_label, levels = income_levels_ordered)
  )

# Reshape from long to wide: columns = Received / Preferred proportions
df_wide <- df_long %>%
  select(income_group_label, format, type, proportion) %>%
  pivot_wider(names_from = type, values_from = proportion)

# Ensure levels are consistent for plotting
df_wide <- df_wide %>%
  mutate(format = fct_rev(factor(format))) %>%
  mutate(income_group_label = factor(income_group_label, levels = income_levels_ordered))

# Plot: Received vs Preferred
ggplot(df_wide, aes(x = Received, y = Preferred, color = format)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 3) +
  facet_wrap(~ income_group_label, ncol = 2) +
  scale_x_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(
    x = "% Received", 
    y = "% Preferred", 
    color = "Format Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    strip.text = element_text(size = 13)
  )

# Save the plot
ggsave(file.path(file_path_plots, "Received_vs_Preferred_by_income_scatter.png"), width = 14, height = 10)
```

Figure 1

```{r echo=FALSE}

communicated_how_region <- summarize_columns_count(
  df = df,
  group_col = "income_group", 
  summary_cols = c("Q16_1", "Q16_2", "Q16_3", "Q16_6", 
                   "Q16_7", "Q16_8", "Q16_9", "Q16_10", 
                   "Q16_11", "Q16_12", "Q16_13", "Q16_14"), 
  section = "metrics_section_answered")

names(communicated_how_region) <- gsub("\\(.*?\\)|\\[.*?\\]", "", names(communicated_how_region))
names(communicated_how_region) <- gsub("\\s$", "", names(communicated_how_region)) 
names(communicated_how_region) <- gsub("\\)$", "", names(communicated_how_region))

# Reshape the data from wide to long format
df_long <- communicated_how_region %>% 
  pivot_longer(cols = c(3:14),
               names_to = "way_of_communication",
               values_to = "count") %>%
  group_by(income_group) %>%
  mutate(proportion = count / n) %>%
  ungroup()

df_long <- df_long %>%
  mutate(way_of_communication = str_wrap(way_of_communication, width = 30))

df_long$income_group_label <- paste0(df_long$income_group, " (n = ", df_long$n, ")")

# Step 1–2: Compute overall top 3 and bottom 3
top_bottom_labels <- df_long %>%
  group_by(way_of_communication) %>%
  summarise(avg_prop = sum(count, na.rm = TRUE)) %>%
  arrange(desc(avg_prop)) %>%
  mutate(color_group = case_when(
    row_number() <= 3 ~ "Top 3 Overall",
    row_number() > n() - 3 ~ "Bottom 3 Overall",
    TRUE ~ "Other"
  )) %>%
  select(way_of_communication, color_group)

# Step 3: Join and label
df_long <- df_long %>%
  left_join(top_bottom_labels, by = "way_of_communication") %>%
  mutate(
    income_group = factor(income_group, levels = c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific")),
    income_group_label = paste0(income_group, " (n = ", n, ")"),
    way_of_communication = reorder_within(way_of_communication, -proportion, income_group)
  ) %>%
  mutate(
    income_group_label = factor(income_group_label, levels = paste0(
      c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"),
      " (n = ", n[match(c("HIC", "UMIC", "LMIC", "LIC", "Non-country-specific"), income_group)], ")"
    ))
  )

# Step 4: Plot
ggplot(df_long, aes(x = way_of_communication, y = proportion, fill = color_group)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Top 3 Overall" = "darkgreen", 
                               "Bottom 3 Overall" = "firebrick", 
                               "Other" = "grey70")) +
  labs(x = "Metric", y = "Proportion", fill = "") +
  facet_wrap(. ~ income_group_label, scales = "free_x", ncol = 2) +
  scale_x_reordered() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom", 
        strip.text = element_text(size = 14),
        plot.margin = margin(t = 10, r = 10, b = 30, l = 20))

ggsave(file.path(file_path_plots, "barplot_metrics_by_income_colored_overall.png"), width = 13.3, height = 13)

```

Figure 3: Confidence plot

```{r}
# Step 1: Summarize proportions
Q24_prop_region <- summarize_columns_count(
  df = df,
  group_col = "income_group",
  summary_cols = paste0("Q24_", 1:9),
  section = "confidence_section_answered"
)

# Step 2: Reshape and clean
df_long <- Q24_prop_region %>%
  pivot_longer(cols = 3:11, names_to = "phase", values_to = "count") %>%
  filter(income_group != "Non-country-specific") %>%
  mutate(
    phase = case_when(
      phase == "Missing" ~ "Other",
      phase == "character(0)" ~ "Other",
      TRUE ~ phase
    ),
    phase = str_wrap(phase, width = 40)
  )

# Step 3: Define income group labels and plotting variables
ordered_income_groups <- c("HIC", "UMIC", "LMIC", "LIC")

group_sizes <- df_long %>%
  select(income_group, n) %>%
  distinct()

ordered_labels <- paste0(
  ordered_income_groups, 
  " (n = ", group_sizes$n[match(ordered_income_groups, group_sizes$income_group)], ")"
)

df_summary <- df_long %>%
  mutate(
    proportion = count / n,
    label = paste0(round(proportion * 100), "%\n(", count, " / ", n, ")"),
    direction = ifelse(income_group %in% c("HIC", "UMIC"), "left", "right"),
    prop_plot = ifelse(direction == "left", -proportion, proportion),
    income_group_label = paste0(income_group, " (n = ", n, ")")
  ) %>%
  mutate(
    income_group_label = factor(income_group_label, levels = ordered_labels)
  )

# Step 4: Order phases by total proportion
phase_order <- df_summary %>%
  group_by(phase) %>%
  summarise(total_prop = sum(proportion, na.rm = TRUE)) %>%
  arrange(desc(total_prop)) %>%
  pull(phase)

df_summary <- df_summary %>%
  mutate(phase = factor(phase, levels = phase_order))

# Step 5: Define consistent color mapping
fill_colors <- setNames(brewer.pal(n = length(ordered_labels), name = "Set2"), ordered_labels)

# Step 6: Plot
ggplot(df_summary, aes(x = prop_plot, y = phase, fill = income_group_label)) +
  geom_col() +
  geom_vline(xintercept = 0, color = "black") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    color = "black", size = 4
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = fill_colors) +
  labs(x = NULL, y = NULL, title = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 16),
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )

# Save plot (optional)
ggsave(file.path(file_path_plots, "confidence_in_modelling_results.png"), width = 16, height = 8)
```

Figure 4: What are barriers to using evidence from epidemic forecasts? Differences between income groups.

```{r echo=FALSE}
# Step 1: Summarize responses by income group
Q25_prop_region <- summarize_columns_count(
  df = df, 
  group_col = "income_group", 
  summary_cols = c("Q25_1", "Q25_10", "Q25_4", "Q25_11", "Q25_5", "Q25_12", "Q25_2", "Q25_3", "Q25_7", "Q25_8", "Q25_9")
)

# Step 2: Reshape and clean
df_long <- Q25_prop_region %>%
  pivot_longer(cols = 3:13, names_to = "phase", values_to = "count") %>%
  filter(income_group != "Non-country-specific") %>%
  mutate(
    phase = ifelse(phase %in% c("Missing", "character(0)"), "Other", phase),
    phase = str_wrap(phase, width = 62)
  )

df_long <- df_long %>%
  filter(count > 0)
# Step 3: Prepare labels and plotting values
ordered_income_groups <- c("HIC", "UMIC", "LMIC", "LIC")

df_summary <- df_long %>%
  mutate(
    proportion = count / n,
    label = paste0(round(proportion * 100), "%\n(", count, " / ", n, ")"),
    direction = ifelse(income_group %in% c("HIC", "UMIC"), "left", "right"),
    prop_plot = ifelse(direction == "left", -proportion, proportion),
    income_group = factor(income_group, levels = ordered_income_groups),
    income_group_label = paste0(income_group, " (n = ", n, ")")
  )

# Step 4: Define label order and phase order
group_sizes <- df_summary %>%
  select(income_group, income_group_label) %>%
  distinct() %>%
  arrange(income_group)

ordered_labels <- group_sizes$income_group_label

df_summary <- df_summary %>%
  mutate(income_group_label = factor(income_group_label, levels = ordered_labels))

phase_order <- df_summary %>%
  group_by(phase) %>%
  summarise(total_prop = sum(proportion, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_prop)) %>%
  pull(phase)

df_summary <- df_summary %>%
  mutate(phase = factor(phase, levels = phase_order)) %>%
  arrange(phase, income_group)

# Step 5: Define consistent color mapping
income_group_colors <- setNames(
  brewer.pal(n = length(ordered_labels), "Set2"),
  ordered_labels
)

# Step 6: Plot
ggplot(df_summary, aes(x = prop_plot, y = phase, fill = income_group_label)) +
  geom_col() +
  geom_vline(xintercept = 0, color = "black") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            color = "black", size = 4) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = income_group_colors) +
  labs(x = NULL, y = NULL, title = "") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 16),
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )

# Step 7: Export
ggsave(file.path(file_path_plots, "barriers_by_income.png"), width = 15, height = 10)
```
